---
title: "Fig6&S6"
author: "TYJ"
date: '2022-05-18'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, background="transparent", fig.align='center' }
knitr::opts_chunk$set(echo = F, include = T, highlight = T, warning = F, fig.align = "c")
options(knitr.kable.NA = '')
library(pheatmap)
library(grid)
library(data.table)
library(GenomicRanges)
library(tidyverse)
source("D:/SUST/code/TanYongjun_code.R")
```

# Four methods

# . upset plot (Overlap among four sets of enhancers)
Steps:
  1. generate list of unique regions. (600bp width). When merge overlapped regions and resize to generate list of  unique CREs, the total number will not match.
  2. generate overlap maptrix (unique regions X Four sets of enhancers) and upset plot.
  3. manually correct numbers in the upset plot.
Attention:
  Overlap between regions may lead to several adjacent CRE (in same dateset) located in the same unique region.
  The "unique regions" list contains redundant peaks. But the rmdup was challenge because width of regions varied and more than two regions may overlapped to form cluster. (比如多个相邻的STARRseq enhancer可能都与另外一个enhancer重叠，如果简单进行merge，会导致upset plot中各类enhancer总数下降)。
  
```{r, cSTARRseq 1.3}
  rm(list = ls())
  source("D:/SUST/code/TanYongjun_code.R")
  enhancer <- fread(file = "./Enhancer_Four_types_Random_PAS_all_20220420.csv.gz") %>%
    dplyr::filter(str_detect(method, "iSTARR|PAS|Random", negate = T),
                  seqnames != "Mt") %>%
    dplyr::mutate(start = startBK, end = endBK) %>%
    dplyr::select(seqnames:end, method) %>%
    dplyr::mutate(method = str_replace_all(method, c("cSTARR-seq" = "Enhancer_STARRseq",
                                                     "IntronicEnhancer" = "Enhancer_Meng2021")),
                  method = toolkit_tyj$Format_En_lev(method) %>% as.character(),
                  ID = str_c(method, seqnames, start, sep = "_"))
  
  # Only keep 941 intronic enhancers
    table(enhancer$method)
    tmp1 <- enhancer %>%
      dplyr::filter(method != "Meng2021")
    tmp2 <- enhancer %>%
      dplyr::filter(method == "Meng2021")
    
    # discard Intronic enhancers overlapped with distal-DHS enhancer.
    tmpDistal <- enhancer %>%
      dplyr::filter(method == "Zhu2015") %>%
      GenomicRanges::makeGRangesFromDataFrame(.) %>%
      GenomicRanges::resize(width = 600, fix = "center")
    ol <- findOverlaps(tmpDistal, 
                       makeGRangesFromDataFrame(tmp2) %>% GenomicRanges::resize(width = 600, fix = "center"))
    set.seed(1234567)
    tmp2 <- tmp2[-unique(ol@to),] %>%
      slice_sample(n = 941, replace = F)
    
    enhancer <- rbind(tmp1, tmp2)
    table(enhancer$method)
    rm(tmp1, tmp2)
  
  # to GR object
  enhancerGR <- enhancer %>%
    split(.$method) %>%
    map(., GenomicRanges::makeGRangesFromDataFrame, keep.extra.columns = T)
  
## Count overlap between enhancers identified by varied methods.
  df <- combn(names(enhancerGR), m = 2) %>% t() %>% as.data.frame()
  df$Num_1 <- df$Num_2 <- df$Num_ov <- NULL
  for (i in 1:nrow(df)) {
    n1 <- df$V1[i]
    n2 <- df$V2[i]
    cat(">>>>  ", n1, "<=>", n2, ";\n")
    tmp1 <- enhancerGR[[n1]]
    tmp2 <- enhancerGR[[n2]]
    ol <- findOverlaps(tmp1, tmp2, minoverlap = 1)
    
    # count
    df$Num_1[i] <- length(tmp1)
    df$Num_2[i] <- length(tmp2)
    df$Num_ov[i] <- length(unique(ol@from)) # number of overlapped regions
  }
  df$Percent_1 <- df$Num_ov / df$Num_1
  df$Percent_2 <- df$Num_ov / df$Num_2
  
  fwrite(df, file = "./R3.enhancer_DistalDHS_H3K4me1/Overlaps_1bp_cenhancers1.3_STARR_DHS_H3K4_941IntronicEn_202210.csv")


## the generate overlap information table: row for enhancer; column for method; value 1 and 0 represent overlapped or not.
  # generate table of unique regions (600bp center of each region or merged region).
  map(enhancerGR, function(x){quantile(width(x), probs = seq(0,1,0.1))})
  
  over_table <- enhancer %>% 
    makeGRangesFromDataFrame() %>%
    GenomicRanges::resize(., width = 600, fix = "center") %>%
    GenomicRanges::reduce() %>%
    GenomicRanges::resize(., width = 600, fix = "center") %>%
    as.data.frame() %>%
    mutate(Enhancer_STARRseq = 0, Enhancer_Zhu2015 = 0, Enhancer_Wang2019 = 0, Enhancer_Meng2021 = 0)
  
  over_tableGR <- makeGRangesFromDataFrame(over_table)
  for (i in 1:length(enhancerGR)) {
    id_tmp <- names(enhancerGR)[i]
    ol <- findOverlaps(over_tableGR, 
                       enhancerGR[[i]]
                       # GenomicRanges::resize(enhancerGR[[i]], width = 1, fix = "center")
                       )
    over_table[unique(ol@from),
               str_detect(colnames(over_table), id_tmp)] <- 1
  }
  
  # clean 
  colnames(over_table) <- str_replace_all(colnames(over_table), "Enhancer_","")
  
  # upsetR
  # discard overlapped enhancers which were repeatedly count in the table.
  library(UpSetR)
  require(ggplot2); require(plyr); require(gridExtra); require(grid);
  
  map_int(enhancerGR, length) %>% sort
  colSums(over_table[,6:9]) %>% sort
  
  # discard overlap between Meng2021 and Zhu2015 due to unique region generation.
  over_table$Meng2021[over_table$Meng2021 == 1 & over_table$Zhu2015 == 1] <- 0
  
  # png("./R3.enhancer_DistalDHS_H3K4me1/UpsetPlot_Four_methods_enhancers_overlap.png",
  #       width = 8, height = 6, units = "cm", res = 600)
  pdf("./R3.enhancer_DistalDHS_H3K4me1/UpsetPlot_Four_methods_enhancers_overlap.pdf",
          width = 9/.pt, height = 6/.pt)
    upset(over_table, 
          sets = unique(enhancer$method) %>% as.character(), 
          order.by = c("freq", "degree"), 
          decreasing = c(TRUE,FALSE),
          set_size.numbers_size = 6,
          set_size.scale_max = 9500,
          set_size.show = T,
          point.size = 0.8, 
          line.size = 0.2,
          sets.bar.color = toolkit_tyj$AtEn_FourMethods_color[c(2,1,3,4)],
          text.scale = 0.8)
  dev.off()
  
  # regions hit by varied kind of enhancers.
  # over_table$Sum = apply(over_table[,6:9], 1, sum)

```


# Distribution (TSS-TTS)----------

### .Enhancer identified by four methods
*Add intronic enhancer 202106*
```{r}
# calculate the relative bins according to gene body.
rm(list = ls())
load(file = "./R4.Distribution_of_enhancers/Enhancer_annotation_ChIPseeker_Peak_up500_down0_IntronicEnhancer_20210605.bin")
# load(file = "./R4.Distribution_of_enhancers/Enhancer_annotation_ChIPseeker_Peak_up500_down0_iSTARR_cSTARR_20210415.bin")
source("D:/SUST/code/TanYongjun_code.R")
anno_table <- anno_list %>% 
  map(., as.data.frame) %>%
  purrr::reduce(rbind) %>%
  toolkit_tyj$relativeBins2TSS(flank.region = 5000, bins_flank = 50, bins_gene = 50, flank_scale = 1) %>%
  dplyr::mutate(method = str_replace_all(method, c("IntronicEnhancer" = "Enhancer_Meng2021", 
                                           "cSTARR-seq" = "Enhancer_STARRseq")))

### Enhancer, PAS, Random------------
# normalized to random
  df <- anno_table %>%
    dplyr::select(method:Type, relativeBins) %>%
    dplyr::mutate(Type = str_replace_all(method, "-", "_")) %>%
    dplyr::filter(!is.na(relativeBins)) %>%
    group_by(relativeBins, Type) %>%
    dplyr::summarise(Num = n()) %>%
    pivot_wider(id_cols = relativeBins, names_from = Type, 
                values_from = Num, values_fill = 0) %>%
    pivot_longer(cols = matches("(Enhancer|PAS|STARR|Random)")) %>%
    pivot_wider(id_cols = c(relativeBins), names_from = name, values_from = value) 
  
  df <- cbind(df$relativeBins, scale(df[,2:ncol(df)]))%>%
    as.data.frame() %>%
    dplyr::rename(relativeBins = V1) %>%
    pivot_longer(cols = matches("(Enhancer|PAS|STARR)")) %>%
    dplyr::mutate(NormedValue = value - Random)
  
  # plot all
  library(ggformula)
  df %>%
    dplyr::filter(str_detect(name, "(iSTARR|PAS)", negate = T)) %>%
    dplyr::mutate(name = toolkit_tyj$Format_En_lev(name) )%>%
    ggplot() + 
      toolkit_tyj$theme_my +
      geom_rect(xmin = 0, xmax = 50, ymin = -3.2, ymax = -2.8, fill = "grey80") +
      geom_text(x = 25, y = -3, label = "Gene body", size = 1.8, color = "grey30") +
      geom_hline(yintercept = 0, linetype = 3, color = "grey40") +
      # geom_line() +
      geom_spline(aes(x = relativeBins, color = name, y = NormedValue, linetype = name)) +
      # scale_color_manual(values = pal_npg(palette = "nrc")(9)[c(1, 2, 3, 4, 6)]) +
      scale_color_npg() +
      scale_linetype_manual(values = c(1,2,2,2,2)) +
      xlab(label = "Distance to TSS/TTS (kb)") + 
      ylab(label = "Relative density") + 
      scale_x_continuous(breaks = c(-50, -25, 0, 10, 20, 30, 40, 50, 75, 100),
                         labels = c("-5", "-2.5", "TSS", "20%", "40%", "60%", "80%", "TTS",
                                    "2.5", "5")) +
      theme(legend.key.size = unit(0.4, units = "cm"), 
          legend.position = c(0.83, 0.22),
          # legend.position = "none",
          legend.title = element_blank(), 
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          legend.key.height = unit(0.25, units = "cm")) 
  
  # toolkit_tyj$SavePlot(filename_prefix = "./R4.Distribution_of_enhancers/Enhancer_distribution_TTS_TSS_STARR_DHS_20210606",
  #                      width = 8, height = 6)

  
# Not normalized
  df <- anno_table %>%
    dplyr::select(method:Type, relativeBins) %>%
    dplyr::mutate(Type = str_replace_all(method, "-", "_")) %>%
    dplyr::filter(!is.na(relativeBins)) %>%
    group_by(relativeBins, Type) %>%
    dplyr::summarise(Num = n()) %>%
    pivot_wider(id_cols = relativeBins, names_from = Type, 
                values_from = Num, values_fill = 0) %>%
    pivot_longer(cols = matches("(Enhancer|PAS|STARR|Random)")) %>%
    pivot_wider(id_cols = c(relativeBins), names_from = name, values_from = value) 
  
  df <- cbind(df$relativeBins, scale(df[,2:ncol(df)]))%>%
    as.data.frame() %>%
    dplyr::rename(relativeBins = V1) %>%
    pivot_longer(cols = matches("(Enhancer|PAS|STARR|Random)")) 
  
  # plot all
  library(ggformula)
  df %>%
    dplyr::filter(str_detect(name, "(iSTARR|PAS)", negate = T)) %>%
    dplyr::mutate(name = toolkit_tyj$Format_En_lev(name)) %>%
    # dplyr::filter(str_detect(name, "Random|Enhancer")) %>%
    ggplot() + 
      toolkit_tyj$theme_my +
      geom_rect(xmin = 0, xmax = 50, ymin = -3.2, ymax = -2.8, fill = "grey80") +
      geom_text(x = 25, y = -3, label = "Gene body", size = 1.8, color = "grey30") +
      # geom_hline(yintercept = 0, linetype = 3, color = "grey40") +
      # geom_line() +
      geom_spline(aes(x = relativeBins, color = name, y = value, linetype = name)) +
      # scale_color_manual(values = pal_npg(palette = "nrc")(9)[c(1, 2, 3, 4, 6)]) +
      scale_color_manual(values = toolkit_tyj$AtEn_FourMethods_color) +
      scale_linetype_manual(values = c(1,1,1,1,2)) +
      xlab(label = "Distance to TSS/TTS (kb)") + 
      ylab(label = "Relative density") + 
      scale_x_continuous(breaks = c(-50, -25, 0, 10, 20, 30, 40, 50, 75, 100),
                         labels = c("-5", "-2.5", "TSS", "20%", "40%", "60%", "80%", "TTS",
                                    "2.5", "5")) +
      theme(legend.key.size = unit(0.4, units = "cm"), 
          legend.position = c(0.83, 0.82),
          # legend.position = "none",
          legend.title = element_blank(), 
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          legend.key.height = unit(0.25, units = "cm")) 
  
  toolkit_tyj$SavePlot(filename_prefix = "./R4.Distribution_of_enhancers/Enhancer_distribution_notNormed_TTS_TSS_Enhancer_fourMethods_20210606",
                       width = 8, height = 6)

```

# Distribution (GAT)-----------
`4.1.3`


# Overlap TE ---------------

`Fig5&S5`

Overlap between TE family/superfamily and CRE

## collect results 202106
```{r, 20210629 }
# load
  rm(list = ls())
  te_info <- fread("./refgenome/TAIR10_Transposable_Elements_20210510_aquired.txt") %>%
    dplyr::mutate(seqnames = str_replace_all(Transposon_Name, "AT(\\d)TE.+", "\\1"), 
                  strand = ifelse(orientation_is_5prime == "TRUE", "+", "-"), 
                  start = Transposon_min_Start, 
                  end = Transposon_max_End) %>%
    dplyr::select(matches("Family")) %>%
    distinct()

  source("D:/SUST/code/TanYongjun_code.R")
  df <- NULL
  for (i in list.files(path = "./R4.Distribution_of_enhancers/TEs/",
                       pattern = "CRE_ol_TEs.+20210629.tsv")) {
    id <- str_replace_all(i, "CRE_ol_TEs_GAT_(.+)20210629.tsv", "\\1")
    df <- fread(str_c("./R4.Distribution_of_enhancers/TEs/", i, sep = "")) %>%
      dplyr::mutate(TE_lev = id,
                    SigLabel = toolkit_tyj$returnAsterisk(qvalue),
                    label = str_c(round(fold, digits = 2), " (", SigLabel, ")"),
                    label1 = str_c(round(fold, digits = 2), "\n", SigLabel),
                    annotation = str_replace_all(annotation, "cSTARR-seq", "Enhancer_STARRseq"),
                    track = str_c(track, " (n=", track_nsegments, ")", sep = ""),
                    track = str_replace_all(track, "TE_TAIR10", "All_TEs")) %>%
      rbind(., df)
  }
  table(df$annotation, df$TE_lev)
  df$track <- str_replace_all(df$track, "_","/")
```

## . Barplot (each CRE)

### . percentage of CRE ov TEs (length)

#### . Four methods

```{r}
 # selecte top TE families.
  top_family <- df %>%
    dplyr::filter(TE_lev == "superFamily", annotation == "Enhancer_STARRseq") %>%
    dplyr::select(track, overlap_size) %>%
    distinct() %>%
    arrange(desc(overlap_size)) %>%
    dplyr::mutate(Track_format = str_replace_all(as.character(track), "\\s\\(.+", ""),
                  Track_format = str_replace_all(Track_format, "_", "/")) %>%
    head(n = 8)

  df_tmp <- df %>%
    dplyr::filter(str_detect(annotation, "(Enhancer_)"), TE_lev == "superFamily") %>%
    dplyr::mutate(observed = observed / annotation_size,
                  expected = expected / annotation_size,
                  track = ifelse(track %in% top_family$track, track, "Other"),
                  Track_format = str_replace_all(as.character(track), "\\s\\(.+", ""),
                  Track_format = str_replace_all(Track_format, "_", "/")) %>%
    group_by(track, annotation, Track_format) %>%
    dplyr::summarise(observed = sum(observed), expected = sum(expected)) %>%
    ungroup() %>%
    dplyr::mutate(expected = ifelse(annotation != "Enhancer_STARRseq", NA, expected)) %>%
    pivot_longer(cols = c(observed, expected), names_to = "Type", values_to = "Rate") %>% 
    na.omit %>%
    dplyr::mutate(Type = ifelse(Type == "expected", "Random", annotation),
                  Track_format = factor(Track_format, levels = c(top_family$Track_format, "Other"))) %>%
    arrange(Track_format) %>%    
    dplyr::mutate(Type = toolkit_tyj$Format_En_lev(Type)) 
  
  df_lab <- df_tmp %>%
    group_by(annotation, Type) %>%
    dplyr::summarise(SumRate = sum(Rate))

  ggplot() + 
      toolkit_tyj$theme_my + 
      geom_col(data = df_tmp,aes(x = Type, y = Rate, fill = Track_format)) + 
      # facet_grid(. ~ annotation) +
      # scale_fill_manual(values = c(pal_npg("nrc")(8), "grey65")) +
      scale_fill_brewer(palette = "Paired") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
            axis.title.x = element_blank(),
            legend.key.size = unit(0.3, units = "cm"),
            legend.margin = margin(0, 0, 0, -8)) + 
      guides(fill = guide_legend(title = "Super Family")) +
      ylab("Percentage") +
      scale_y_continuous(labels = scales::percent) + 
      geom_text(data = df_lab, aes(x = Type, y = SumRate + 0.01, label = percent(SumRate, accuracy = 0.01)),
                size = 5/.pt)
  
  toolkit_tyj$SavePlot(filename_prefix = "./R4.Distribution_of_enhancers/TEs/Percentage_Enhancer_fourmethods_overlap_TE_superfamily_20220521",
                       width = 6, height = 6)
```

### . FC of overlap

#### . Enhancers four methods
```{r}
  top_family <- df %>%
    dplyr::filter(TE_lev == "superFamily", annotation == "Enhancer_STARRseq") %>%
    dplyr::select(track, overlap_size) %>%
    distinct() %>%
    # dplyr::filter(overlap_size > 0) %>%
    arrange(desc(overlap_size)) %>%
    head(., n = 8) 

  df_tmp <- df %>%
    dplyr::filter(str_detect(annotation, "(Enhancer_)"), TE_lev == "superFamily",
                  track %in% top_family$track) %>%
    dplyr::mutate(track = factor(track, top_family$track)) %>%
    arrange(track) %>%
    dplyr::mutate(annotation = str_replace_all(annotation, "Enhancer_", "En_"),
                  annotation = toolkit_tyj$Format_En_lev(annotation))
    # dplyr::mutate(observed = observed / annotation_size,
    #               expected = expected / annotation_size,
    #               track = ifelse(track %in% top_family$track, track, "Other")) %>%
    # group_by(track, annotation) %>%
    # summarise(observed = sum(observed), expected = sum(expected)) %>%
    # ungroup() 
  
  df_tmp %>%
    ggplot(aes(y = track, x = l2fold, fill = l2fold)) + 
      toolkit_tyj$theme_my + 
      geom_col(size = 0.2) + 
      facet_grid(. ~ annotation) +
      scale_fill_gradientn(colors = brewer.pal(11, "RdYlGn")[c(2,6:10)] %>% rev) +
      # scale_fill_brewer(palette = "Paired") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
            axis.title.y = element_blank(),
            legend.key.width  = unit(0.3, units = "cm"),
            legend.margin = margin(0, 0, 0, -5),
            panel.spacing = unit(0.08, "cm")) + 
      guides(fill = guide_colourbar(title = expression(log[10]^(FC)))) +
      scale_y_discrete(labels = function(x){str_replace_all(x, "\\(", "\n\\(")},
                       limits = rev) +
      ylab("Percentage") +
      geom_text(aes(x = ifelse(abs(l2fold) < 2, l2fold + (l2fold^0) * 2, l2fold/2),
                    y = track, label = SigLabel), size = 5/.pt) +
      xlab(expression(log[10]^(Fold~change)))
      
  
  toolkit_tyj$SavePlot(filename_prefix = "./R4.Distribution_of_enhancers/TEs/Barplot_fold_change_TEsuperfamily_ol_four_enhancers_1_20220521",
                       width = 8, height = 6)
```



# Overlap (En vs Conserved elements)

```{r}
  rm(list = ls())
  source("D:/SUST/code/TanYongjun_code.R")
  results_sig <- fread("./R8.conservation/ConservedElements_ol_GAT_results_all_20210630.tsv") %>%
    dplyr::mutate(percent_overlap_size_annotation = percent_overlap_size_annotation / 100,
                  annotation = toolkit_tyj$Format_En_lev(annotation),
                  qvalue = p.adjust(pvalue, method = "BH"),
                  label = str_c(scales::percent(percent_overlap_size_annotation, accuracy = 0.01),
                                "\nFC=", format(fold, digits = 2), "\n", toolkit_tyj$returnAsterisk(qvalue)),
                  observed = observed / annotation_size,
                  expected  = expected / annotation_size)
  
  # Enhancers identified by four methods
  results_sig %>%
    ggplot(aes(x = annotation, y = percent_overlap_size_annotation)) + 
      geom_col(aes(fill = annotation), alpha = 3/4) + 
      toolkit_tyj$theme_my + 
      theme(axis.title.x = element_blank(),
            axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
            legend.title = element_blank(),
            legend.margin = margin(-8,0,0,-8), 
            legend.key.size = unit(0.3, "cm"),
            text = element_text(face = "plain")) +
      scale_y_continuous(labels = scales::percent) + 
      scale_fill_manual(values = toolkit_tyj$AtEn_FourMethods_color) +
      geom_text(aes(y = percent_overlap_size_annotation / 2, label = label), size = 5/.pt) + 
      ylab("Enhancers overlaped with conserved elements")
  
  toolkit_tyj$SavePlot(filename_prefix = "./R8.conservation/ConservedElements_ol_GAT_results_BarPlot_20210514",
                       width = 6, height = 6)
  
  # only Enhancer_STARRseq
  results_sig %>%
    dplyr::filter(annotation == "En_STARRseq") %>%
    pivot_longer(cols = c(expected, observed), names_to = "Type", values_to = "percent") %>%
    dplyr::mutate(Type = str_replace_all(Type, c("observed" = "Observed", 
                                                 "expected" = "Expected")),
                  Type = factor(Type, levels = c("Observed", "Expected"))) %>%
    ggplot(aes(x = Type, y = percent)) + 
      geom_col(aes(fill = Type), alpha = 2/3) + 
      toolkit_tyj$theme_my + 
      theme(axis.title.x = element_blank(),
            axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
      scale_y_continuous(labels = scales::percent) + 
      geom_text(aes(y = percent / 2, label = percent(percent, accuracy = 0.01)), size = 5/.pt) + 
      scale_fill_manual(values = toolkit_tyj$AtEn_color[c(1, 5)]) +
      ylab("Enhancers overlaped with conserved elements") +
      geom_signif(annotations = "FC=1.42\np<0.001",
                  y_position = 0.4, xmin = 1, xmax = 2,textsize  = 6/.pt, size = 0.2) + 
      theme(legend.position = "none") + 
      coord_cartesian(ylim = c(0, 0.46))
  
  toolkit_tyj$SavePlot(filename_prefix = "./R8.conservation/ConservedElements_ol_Enhancer_STARRseq_GAT_results_BarPlot_20220520",
                       width = 5, height = 6)
  
  
```


# TFs
 Related code was contained in `Fig3&S3.rmd`
 

# Revise (add super enhancers and enhancers identified by Yan et al.)





